Minimax-optimal Laplacian Eigenmaps regression over Sobolev Spaces with Neighborhood Graphs

### 2021-03-02 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
More progress made on the draft of Laplacian Eigenmaps paper.

\textbf{(1) Sobolev norms, semi-norms, and balls.} 
The Sobolev class $H^s(P)$ consists of all those functions $f \in \Leb^2(P)$ which have weak partial derivatives of all orders $j = 1,\ldots,s$, all of which are bounded in $\Leb^2(P)$ norm. The Sobolev norm $\|f\|_{H^s(P)}$ is defined according to
\begin{equation*}
\|f\|_{H^s(P)}^2 := \|f\|_{\Leb^2(P)}^2 + \sum_{j = 1}^{s} \|D^jf\|_{\Leb^2(P)}^2,
\end{equation*}
where $D^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$, and we use the multiindex notation $D^{\alpha}f = \partial^j f/(\partial^{\alpha_1}x^1\ldots \partial^{\alpha_d}x^d)$ for multiindex $\alpha$. The Sobolev ball of radius $M$ consists of all those functions $f \in H^s(P)$ with Sobolev norm no greater than $M$,
\begin{equation*}
H^s(P;M) := \{f \in H^s(P): \|f\|_{H^s(P)} \leq M\}
\end{equation*}

@@

\subsection{Our contributions}
The following items summarize our major results. In all of them we assume~\ref{asmp:domain} and~\ref{asmp:density} hold, but not that any of $P$, $\mc{X}$, $\Delta_P$ or $\Delta$ are known. Put $r := s - 1$.
\begin{itemize}
	\item \textbf{Minimax optimal estimation.} For any $f_0 \in H^1(\Xset;M)$, with high probability $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2 + d)}n^{-2/(2 + d)}$.
	\item \textbf{Minimax optimal testing.}
	A test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(2 + d)}n^{-4/(4 + d)}$ and $d < 4$.
	\item \textbf{Higher-order smoothness conditions.} When the regression function $f_0 \in H_0^s(\Xset;M)$ and additionally the density $p \in C^{r}(\mc{X};M)$, with high probability the error $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim M^{2d/(2s + d)}n^{-2s/(2s + d)}$. Additionally, a test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(4s + d)}n^{-4s/(4s + d)}$ and $d < 4$.
	\item \textbf{Manifold adaptivity.}
	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$ and $s \leq 2$, each of the aforementioned rates hold with $d$ replaced by $m$.
\end{itemize}
The above estimation results are stated with respect to in-sample mean squared error, meaning the loss is measured in the empirical squared-norm $\|\cdot\|_n^2$.\footnote{To state the obvious: this is very different than measuring the \emph{training error} $\|Y - \wh{f}\|_n^2$.} This is natural, as Laplacian eigenmaps is defined only at the observed design points $X_1,\ldots,X_n$. That being said, often one would like an upper bound on the prediction risk, or equivalently the out-of-sample loss $\|\cdot\|_P^2$. We therefore also introduce a simple, intuitive way of extending $\wh{f}$ to a function defined over all of $\mc{X}$, and show that this extension is minimax rate-optimal with respect to error measured in $L^2(P)$ norm.

\subsection{Related Work}
\begin{itemize}
	\item \textcolor{red}{(TODO)} Summarize what is known on fixed graphs.
	\item \textcolor{red}{Belkin, Slepcev} show that for a fixed $k$, the eigenvector $v_k$ converges to the eigenfunction $u_k$, in the sense that \textcolor{red}{(...)}. \textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$. This justifies our previous discussion, in which we treat Laplacian eigenmaps as an approximation of spectral series methods. However, we emphasize that the minimax optimality of $\wh{f}$ or a test based on $\wh{T}$ does not follow straightforwardly from such results. Instead our analysis proceeds along different lines, which we detail in  Section~\ref{sec:analysis}.
	\item There has been limited analysis of the estimator $\wh{f}$. \cite{zhou2011} consider a similar estimator, but in the semi-supervised setting where the number of unlabeled points grows to infinity. In this case the estimator reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$. \cite{lee2016} considers a similar estimator---using a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but get suboptimal rates for the supervised problem. To the best of knowledge, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$, and study a regularized estimator estimator, which minimizes the penalized loss
	\begin{equation*}
	\|Y - f\|_n^2 + \lambda \dotp{L_{n,\varepsilon}f}{f}_n.
	\end{equation*}
	\textcolor{blue}{(Green 2021)} show that the resulting estimator is (nearly)-minimax optimal, but only for $s = 1$ and $d \leq 4$. In contrast, the Laplacian eigenmaps estimator is optimal for all $s$ and $d$. 
\end{itemize}

@@

\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$. 
\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{3}
	\item 
	\label{asmp:parameters_estimation_fo} 
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
	\begin{equation}\\
	\label{eqn:radius_fo} 
	C_0 \biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \varepsilon \leq \min\{i_0,K^{-1/d}\},
	\end{equation}
	and 
	\begin{equation}
	\label{eqn:eigenvector_estimation_fo} 
	c_1 (M^2 n)^{2d/(2 + d)}\leq K \leq C_1 (M^2 n)^{2d/(2 + d)}.
	\end{equation}
\end{enumerate}
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation_fo}
	Suppose $f_0 \in H^1(P,M)$. If the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, then the following statement holds for any $\delta \in (0,1)$: with probability at least $1 - \delta - \textcolor{red}{(???)}$,
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_estimation_fo}
	\|\wh{f} - f_0\|_n^2 \leq \frac{C}{\delta}M^{2d/(2 + d)}n^{-2/(2 + d)}
	\end{equation}
\end{theorem}
Some remarks:
\begin{itemize}
	\item \textbf{(1.1) How to tune hyper-parameters.} We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as $f_0$ and $P$. In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation. However, because the estimator $\wh{f}$ is defined only in-sample, such approaches cannot be straightforwardly applied to select the graph radius $\varepsilon$, or number of eigenvectors $K$. We discuss out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
	\item \textbf{(1.2) The dependence on failure probability can be improved.} The upper bound on in-sample risk given by Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} depends on the pre-factor $C/\delta$, and holds with probability $1 - \delta - \textcolor{red}{(?)}$. When $f_0 \in C^1(\mc{X};M)$, we can improve these dependencies, changing the pre-factor in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} to $C(1 + 1/\delta)$, and showing that the bound holds with probability at least $1 - C\delta^2/n$. 
\end{itemize}

@@

\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{4}
	\item 
	\label{asmp:parameters_testing_fo}
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_fo}. Additionally,
	\begin{equation}
	\label{eqn:eigenvector_testing_fo}
	c_2 (M^2 n)^{2d/(4 + d)}\leq K \leq C_2 (M^2 n)^{2d/(4 + d)}.
	\end{equation}
\end{enumerate}
Let $p(d) = \min\{1,2/d\}$.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_fo}
	Suppose $f_0 \in H^1(\mc{X};M)$. If the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_fo}, then the following statement holds for any $\delta \in (0,1)$: if $f_0$ satisfies
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
	\|f_0\|_P^2 \geq \frac{C}{\delta} \max\biggl\{\frac{1}{\delta}M^{2d/(4 + d)}n^{-4/(4 + d)}, M n^{-p(d)}\biggr\},
	\end{equation}
	then $\varphi$ has Type II error upper bounded by 
	\begin{equation}
	\Pbb_{f_0}(\varphi = 0) \leq \delta + \textcolor{red}{(...)}
	\end{equation}
\end{theorem}
Some remarks:
\begin{itemize}
	\item \textbf{(2.1) When $d \geq 4$, the testing problem is different.} When $d \geq 4$, it is not the case that $\Leb^4(\mc{X})$ continuously embeds into $H^1(\mc{X})$, and testing rates over $H^1(\mc{X})$ are not known. On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then simply taking $T = \|Y\|_n^2$ achieves the optimal rates.
	\item \textbf{(2.2) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} also apply to testing.}
\end{itemize}

\subsection{Higher-order Sobolev classes}
\label{sec:higher_order_sobolev_classes}
In this section, we now assume that the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X})$. We point out that on a graph there is no natural analogue to higher-order derivatives. For this reason it is not at all obvious that graph-based methods \emph{can} adapt to higher-order smoothness. Remarkably, we show that this is in fact the case. Under appropriate choices of $\varepsilon$ and $K$, the estimator $\wh{f}$ has (in-sample) optimal error (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}$. 
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation_ho}
	Under appropriate conditions, with probability at least $1 - \delta$,
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_estimation_ho}
	\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}
	\end{equation}
\end{theorem}
Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a rate-optimal estimator over Sobolev classes, for all scalings of $s$ and $d$. Some other remarks:
\begin{itemize}
	\item \textbf{(1.1) Optimal rates, no RKHS.} It is worth pointing out that we do not require that the regularity of the Sobolev space satisfy $2s > d$, a condition often seen in the literature. In the ``sub-critical'' regime $2s < d$, the Sobolev space $H^s(\mc{X})$ is quite irregular. For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space. As a result, one has to be a bit careful. Certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $X$ are assumed to be fixed. It also takes certain estimators ``off the table'', most notably RKHS-based methods such as smoothing splines. However, at these scales spectral series estimators still obtain the optimal rates for random design regression measured in $\Leb^2(P)$ norm. Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$ norm.
	\item \textbf{(1.2) The zero-trace condition.} The zero-trace condition can likely be weakened. However, requiring some boundary condition on $f_0$ and its first through $(k -1 )$th-order derivatives is unavoidable. \textcolor{red}{Explain why: graphs are implicitly boundaryless objects, and the eigenvectors of $L$ converge to those eigenvectors of $\Delta_P$ which satisfy boundary conditions. Repeat that this is a standard assumption made when analyzing spectral series estimators}.
	\item \textbf{(1.3) The bounded norm and smooth density conditions.} We require that $f_0$ be bounded in the Sobolev norm $\|f_0\|_{H^s(\mc{X})}$---as opposed to the semi-norm $|f_0|_{H^s(\mc{X})}$--- and that $p \in C^{s - 1}(\mc{X})$. Both requirements are related and fundamental, as we shall now explain. As previously mentioned, $L$ converges to a weighted-version of the Laplace operator $\Delta_P$ (also known as a Fokker-Planck) operator. The operator $\Delta_P$ includes both first and second-order derivatives,
	\begin{equation}
	\label{eqn:fokker_planck_1}
	\Delta_Pf= \frac{-1}{p} \mathrm{div}(p^2 \nabla f) = \Delta f - 2\frac{\nabla p^{\top} \nabla f}{p},
	\end{equation}
	and therefore a function $f \in H_0^s(\mc{X})$ has finite energy $\|\Delta_P f\|_{P}^2$ only if the first-order partial derivatives of $f$ and $p$ are bounded. \textcolor{red}{This discussion---and in particular the last sentence---are obviously not quite correct. This is in part because you realize now that the requirements are not in fact fundamental. Instead, the requirement should merely be that $\|\Delta_P^kf\|_{\Leb^2(P)}^2$ is bounded---here $k = s/2$ and we assume $s$ is even---and the upper bounds should be in terms of this semi-norm.}.
	\item \textbf{(1.4) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} continue to apply.}
\end{itemize}

\textbf{(1.4) Rates for testing with higher-order smoothness assumptions.} For testing, we can get the optimal rate $n^{-4s/(4s + d)}$ when $d \leq 4$. When $4 < d \leq 4s$, we get the rate~\textcolor{red}{(...)} Recall that the rate $n^{-4s/(4s + d)}$ is optimal whenever $d \leq 4s$. Therefore there are scalings of $s$ and $d$ for which we do not get the optimal rate. We do not know whether this due to looseness in our proof techniques, or a problem with the test itself.

@@

\textbf{(1) Bias and variance terms, for estimation and testing}. We analyze the estimation error of $\wh{f}$ by first condition on the design points $X$ and computing \emph{design-dependent} bias and variance terms. For estimation, we have that with high probability, conditional on the design $X$,
\begin{equation*}
\|\wh{f} - f_0\|_n^2 \leq C\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{\kappa}^s} + \frac{K}{n}
\end{equation*}
For testing, we have~\textcolor{red}{(...)}.

The semi-norm $\dotp{L^s f_0}{f_0}_n$ and eigenvalue $\lambda_{K}$ are random variables that depend the design $X$. It remains to establish suitable upper and lower bounds on these quantities.

>>>>

\textbf{(1) Sobolev norms, semi-norms, and balls.} 
The Sobolev class $H^s(P)$ consists of all those functions $f \in \Leb^2(P)$ which have weak partial derivatives of all orders $j = 1,\ldots,s$, all of which are bounded in $\Leb^2(P)$ norm. The Sobolev norm $\|f\|_{H^s(P)}$ is defined according to
\begin{equation*}
\|f\|_{H^s(P)}^2 := \|f\|_{\Leb^2(P)}^2 + \sum_{j = 1}^{s} \|D^jf\|_{\Leb^2(P)}^2,
\end{equation*}
where $D^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$, and we use the multiindex notation $D^{\alpha}f = \partial^j f/(\partial^{\alpha_1}x^1\ldots \partial^{\alpha_d}x^d)$ for multiindex $\alpha$. The Sobolev ball of radius $M$ consists of all those functions $f \in H^s(P)$ with Sobolev norm no greater than $M$,
\begin{equation*}
H^s(P;M) := \bigl\{f \in H^s(P): \|f\|_{H^s(P)} \leq M\bigr\}
\end{equation*}

@@

\subsection{Our contributions}
The following items summarize our major results. In all of them we assume~\ref{asmp:domain} and~\ref{asmp:density} hold, but not that any of $P$, $\mc{X}$, $\Delta_P$ or $\Delta$ are known. Put $r := s - 1$.
\begin{itemize}
	\item \textbf{Minimax optimal estimation.} For any $f_0 \in H^1(\Xset;M)$, with high probability $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2 + d)}n^{-2/(2 + d)}$.
	\item \textbf{Minimax optimal testing.}
	A test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(2 + d)}n^{-4/(4 + d)}$ and $d < 4$.
	\item \textbf{Higher-order smoothness conditions.} When the regression function $f_0 \in H_0^s(\Xset;M)$ and additionally the density $p \in C^{r}(\mc{X};M)$, with high probability the error $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2s + d)}n^{-2s/(2s + d)}$. Additionally, a test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(4s + d)}n^{-4s/(4s + d)}$ and $d < 4$.
	\item \textbf{Manifold adaptivity.}
	If $\mc{X} \subset \Rd$ is a manifold of dimension $m < d$, and if $H^s(\mc{X})$ for $s \leq 2$, then each of the aforementioned rates hold with $d$ replaced by $m$.
\end{itemize}
The above estimation results are stated with respect to in-sample mean squared error, meaning the loss is measured in the empirical squared-norm $\|\cdot\|_n^2$.\footnote{To state the obvious: this is very different than measuring the \emph{training error} $\|Y - \wh{f}\|_n^2$.} This is natural, as Laplacian eigenmaps is defined only at the observed design points $X_1,\ldots,X_n$. That being said, often one would like an upper bound on the prediction risk, or equivalently the out-of-sample loss $\|\cdot\|_P^2$. We therefore also introduce a simple, intuitive way of extending $\wh{f}$ to a function defined over all of $\mc{X}$, and show that this extension is minimax rate-optimal with respect to error measured in $L^2(P)$ norm.

\subsection{Related Work}
\begin{itemize}
	\item \textcolor{red}{(TODO)} Summarize what is known on fixed graphs.
	\item \textcolor{red}{Belkin, Slepcev} show that for a fixed $k$, the eigenvector $v_k$ converges to the eigenfunction $u_k$, in the sense that \textcolor{red}{(...)}. \textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$. This justifies our previous discussion, in which we treat Laplacian eigenmaps as an approximation of spectral series methods. However, we emphasize that the minimax optimality of $\wh{f}$ or a test based on $\wh{T}$ does not follow straightforwardly from such results. Instead our analysis proceeds along different lines, which we detail in  Section~\ref{sec:analysis}.
	\item There has been limited analysis of the estimator $\wh{f}$. \cite{zhou2011} consider a similar estimator, but in the semi-supervised setting where the number of unlabeled points grows to infinity. In this case the estimator reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$. \cite{lee2016} consider a similar estimator---using eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but get suboptimal rates for the supervised problem. As far as we know, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$, and study a regularized estimator estimator, which minimizes the penalized loss
	\begin{equation*}
	\|Y - f\|_n^2 + \lambda \dotp{L_{n,\varepsilon}f}{f}_n.
	\end{equation*}
	\textcolor{blue}{(Green 2021)} show that the resulting estimator is (nearly)-minimax optimal, but only for $s = 1$ and $d \leq 4$. In contrast, the Laplacian eigenmaps estimator is optimal for all $s$ and $d$. 
\end{itemize}

@@

\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$. 
\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{3}
	\item 
	\label{asmp:parameters_estimation_fo} 
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
	\begin{equation}\\
	\label{eqn:radius_fo} 
	C_0 \biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \varepsilon \leq \min\{i_0,K^{-1/d}\},
	\end{equation}
	and 
	\begin{equation}
	\label{eqn:eigenvector_estimation_fo} 
	c_1 (M^2 n)^{2d/(2 + d)}\leq K \leq C_1 (M^2 n)^{2d/(2 + d)}.
	\end{equation}
\end{enumerate}
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation_fo}
	Suppose $f_0 \in H^1(P,M)$. If the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, then the following statement holds for any $\delta \in (0,1)$: with probability at least $1 - \delta - \textcolor{red}{(???)}$,
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_estimation_fo}
	\|\wh{f} - f_0\|_n^2 \leq \frac{C}{\delta}M^{2d/(2 + d)}n^{-2/(2 + d)}
	\end{equation}
\end{theorem}
Some remarks:
\begin{itemize}
	\item Regarding the restrictions on hyperparameters: the lower bound on $\varepsilon$ is the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected. On the other hand, as we will see in Section~\ref{sec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K \asymp \lambda_K(\Delta_P) \asymp K^{2/d}$. Finally, the restriction $K \asymp (M^2n)^{2d/(2 + d)}$ is chosen  to optimally trade-off bias and variance. Together, these restrictions suggest that the 
	\item \textbf{(1.1) How to tune hyper-parameters.} We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$. In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation. However, because the estimator $\wh{f}$ is defined only in-sample, such approaches cannot be straightforwardly applied to select the graph radius $\varepsilon$, or number of eigenvectors $K$. We return to this issue upon considering out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
	\item \textbf{(1.2) The dependence on failure probability can be improved.} The upper bound on in-sample risk given by Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} depends on the pre-factor $C/\delta$, and holds with probability $1 - \delta - \textcolor{red}{(?)}$. When $f_0 \in C^1(\mc{X};M)$, we can improve these dependencies, changing the pre-factor in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} to $C(1 + 1/\delta)$, and showing that the bound holds with probability at least $1 - C\delta^2/n$. 
\end{itemize}

@@

\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{4}
	\item 
	\label{asmp:parameters_testing_fo}
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_fo}. Additionally,
	\begin{equation}
	\label{eqn:eigenvector_testing_fo}
	c_2 (M^2 n)^{2d/(4 + d)}\leq K \leq C_2 (M^2 n)^{2d/(4 + d)}.
	\end{equation}
\end{enumerate}
Let $p(d) = \min\{1,2/d\}$.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_fo}
	Suppose $f_0 \in H^1(\mc{X};M)$. If the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_fo}, then the following statement holds for any $\delta \in (0,1)$: if $f_0$ satisfies
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
	\|f_0\|_P^2 \geq \frac{C}{\delta} \max\biggl\{\frac{1}{\delta}M^{2d/(4 + d)}n^{-4/(4 + d)}, M n^{-p(d)}\biggr\},
	\end{equation}
	then $\varphi$ has Type II error upper bounded by 
	\begin{equation}
	\Pbb_{f_0}(\varphi = 0) \leq \delta + \textcolor{red}{(...)}
	\end{equation}
\end{theorem}
Some remarks:
\begin{itemize}
	\item \textbf{(2.1) When $d \geq 4$, the testing problem is different.} When $d \geq 4$, it is not the case that $H^1(\mc{X})$ continuously embeds into $\Leb^4(\mc{X})$, and as far we know the optimal testing rates over $H^1(\mc{X})$ are not known. On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then simply taking $T = \|Y\|_n^2$ achieves the optimal rates.
	\item \textbf{(2.2) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} also apply to testing.}
\end{itemize}

\subsection{Higher-order Sobolev classes}
\label{sec:higher_order_sobolev_classes}
In this section, we now assume that the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X})$. We point out that on a graph there is no natural analogue to higher-order derivatives. For this reason it is not at all obvious that graph-based methods \emph{can} adapt to higher-order smoothness. Remarkably, we show that this is in fact the case. Under appropriate choices of $\varepsilon$ and $K$, the estimator $\wh{f}$ has (in-sample) optimal error (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}$.

\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{5}
	\item 
	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfies
	\begin{equation*}
	C_0 \cdot \max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, n^{-1/(2r + d)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/d}\}
	\end{equation*}
	and
	\begin{equation*}
	K \asymp (M^2n)^{2d/(2s + d)}
	\end{equation*}
\end{enumerate}

\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation_ho}
	Under appropriate conditions, with probability at least $1 - \delta$,
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_estimation_ho}
	\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}
	\end{equation}
\end{theorem}
Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a rate-optimal estimator over Sobolev classes, for all scalings of $s$ and $d$. Some other remarks:
\begin{itemize}
	\item \textbf{(1.1) Optimal rates, no RKHS.} It is worth pointing out that we do not require that the regularity of the Sobolev space satisfy $2s > d$, a condition often seen in the literature. In the ``sub-critical'' regime $2s < d$, the Sobolev space $H^s(\mc{X})$ is quite irregular. For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space. As a result, certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $X$ are assumed to be fixed. Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as smoothing splines. However, at these scales spectral series estimators still obtain the optimal rates for random design regression measured in $\Leb^2(P)$ norm. Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$ norm.
	\item \textbf{(1.2) The zero-trace condition.} The zero-trace condition can likely be weakened. However, requiring some boundary condition on $f_0$ and its first through $(k -1 )$th-order derivatives is unavoidable. \textcolor{red}{Explain why: graphs are implicitly boundaryless objects, and the eigenvectors of $L$ converge to those eigenvectors of $\Delta_P$ which satisfy boundary conditions. Repeat that this is a standard assumption made when analyzing spectral series estimators}.
	\item \textbf{(1.3) The bounded norm and smooth density conditions.} We require that $f_0$ be bounded in the Sobolev norm $\|f_0\|_{H^s(\mc{X})}$---as opposed to the semi-norm $|f_0|_{H^s(\mc{X})}$--- and that $p \in C^{s - 1}(\mc{X})$. Both requirements are related and fundamental, as we shall now explain. As previously mentioned, $L$ converges to a weighted-version of the Laplace operator $\Delta_P$ (also known as the Fokker-Planck) operator. The operator $\Delta_P$ includes both first and second-order derivatives,
	\begin{equation}
	\label{eqn:fokker_planck_1}
	\Delta_Pf= \frac{-1}{p} \mathrm{div}(p^2 \nabla f) = \Delta f - 2\frac{\nabla p^{\top} \nabla f}{p},
	\end{equation}
	and therefore a function $f \in H_0^s(\mc{X})$ has finite energy $\|\Delta_P f\|_{P}^2$ only if the first-order partial derivatives of $f$ and $p$ are bounded. \textcolor{red}{This discussion---and in particular the last sentence---are obviously not quite correct. This is in part because you realize now that the requirements are not in fact fundamental. Instead, the requirement should merely be that $\|\Delta_P^kf\|_{\Leb^2(P)}^2$ is bounded---here $k = s/2$ and we assume $s$ is even---and the upper bounds should be in terms of this semi-norm.}.
	\item \textbf{(1.4) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} continue to apply.}
\end{itemize}

\textbf{(1.4) Rates for testing with higher-order smoothness assumptions.} For testing, we can get the optimal rate $n^{-4s/(4s + d)}$ when $d \leq 4$. When $4 < d \leq 4s$, we get the rate~\textcolor{red}{(...)} Recall that the rate $n^{-4s/(4s + d)}$ is optimal whenever $d \leq 4s$. Therefore there are scalings of $s$ and $d$ for which we do not get the optimal rate. We do not know whether this is due to looseness in our proof techniques, or a problem with the test itself, and leave this for future work.

@@

\textbf{(1) Bias and variance terms, for estimation and testing}. We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points $X$ and deriving \emph{design-dependent} bias and variance terms. For estimation, we have that with high probability, conditional on the design $X$,
\begin{equation*}
\|\wh{f} - f_0\|_n^2 \leq C\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{\kappa}^s} + \frac{K}{n}
\end{equation*}
For testing, we have~\textcolor{red}{(...)}.

The semi-norm $\dotp{L^s f_0}{f_0}_n$ and eigenvalue $\lambda_{K}$ are random variables that depend the random design points $X_1,\ldots,X_n$. It remains to establish suitable upper and lower bounds on these quantities.

\textbf{(2) Upper bound on the semi-norm. } Proposition~\ref{prop:graph_seminorm} gives a sufficient upper bound on $\dotp{L^s f_0}{f_0}_n$. Take $\beta := s - 1$. 
\begin{proposition}
	\label{prop:graph_seminorm} 
	Suppose $f \in H_0^s(\mc{X})$, that $p \in C^{\beta}(\mc{X})$, and that $r \gtrsim n^{-1/d}$. Then with probability at least $1 - \delta$, it holds that 
	\begin{equation}
	\label{eqn:graph_seminorm_1}
	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}} + \textcolor{red}{(?)}
	\end{equation}
	Therefore, if additionally $r \gtrsim n^{-1/(2\beta + d)}$, then 
	\begin{equation}
	\label{eqn:graph_seminorm_2}
	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}}
	\end{equation}
	When $s = 1$, the results hold for $f \in H^1(\mc{X})$ and any density $p$ that satisfies Assumption~\ref{asmp:density}.
\end{proposition}
The proof of Proposition~\ref{prop:graph_seminorm} involves two steps. In the first step, we show that with probability at least $1 - \delta$, the graph semi-norm $\dotp{L^s f}{f}_n$ is upper bounded by a non-local continuum energy $E_{\varepsilon}^{(s)}(f)$, defined as
\begin{equation*}
E_{\varepsilon}^{(s)}(f) := \dotp{L_{P,\varepsilon}^sf}{f}_{P}
\end{equation*}
where $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$, 
\begin{equation*}
L_{P,\varepsilon}f(x) := \int_{\mc{X}}\bigl(f(z) - f(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(x).
\end{equation*}
In the second step, we show that $E_{\varepsilon}^{(s)}(f)$ is in turn upper bounded by a constant times $D_s(f) := \|f\|_{H^s(P)}^2$. Together, these steps establish Proposition~\ref{prop:graph_seminorm}.
